Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions

### 2024-07-25 latex/refs.tex
Tidy up a bit

\begin{description}
  \item[a] the Introduction is to be made accessible to the whole BA
    readership;
  \item[b] the motivation and the practical statistical implications are to be
    developed and discussed;
  \item[c] the contribution has to be better contextualized with
    respect to the literature in both inverse problems and D-optimal
    designs;
  \item[d] numerical evidence and code, mentioned in the paper
    but actually not provided, need to be made available;
  \item[e] the mathematical derivations have to be watertight and some
    clarifications are needed in this respect (see Report of Referee
    3).

\end{description}

@@

@@

However, I feel changing the assumption from independent to correlated
measurements is quite artificial: people tweak a data generation
process assumption to solve an optimal design problem, and the author
just takes the assumption as it is. Since the author proves the
clusterization issue under the independent measurement assumption, I
expect the author to go deeper and explain why an ordinary data
generation assumption + a common optimal design leads to a
“suboptimal” design? Not just by the mathematical proof, but from an
information theory point of view. For example, does the redundancy
points mean additional points does not reduce the variance of the
estimation at all? If that’s the case, how does correlated measurement
assumption “help” with it?

\answer: Thanks for your questions. I am not a proponent of using
correlated noise, I just wanted to show the usefullness of my model by
rigorously proving how correlated noise mitigates clusterization. I
have added several lines clearing this issue.

@@

\answer:

@@

If one looks at, for example, Fedorov (1996) and Nyberg et
al (2012), it is an undesirable property because they are interested
in sensor location, and two or more sensors cannot be in the same
location. However, there are many experiments where repeated
measurements are possible, and are therefore optimal (if the model is
correct, which it won’t be). An example, is a chemical engineering
experiment where one wishes to learn the relationship between a series
of variables and yield of a chemical reaction. The experiment will
involve specifying the variables, and measuring the yield (the
observation) after a specified period of time. This process is then
duplicated with a potentially different set of variables.  It is
perfectly possible to have repeated variables.

\answer: again

@@

@@

@@

\subsection{Minor comments}
\begin{itemize}
\item The author uses two different forms of asterisk: one for
  optimality of the design (e.g. Definition 3) and one for the
  adjoint. Suggest changing the one for optimality.
\item I do not like the use of the term measurements as used in the
  paper. The measurements are the $\data$’s since these are a result of
  measuring a physical quantity. However, the paper refers to the
  quantities that are controlled as the measurements.
\item Proof of Proposition 12 uses $\eps$ as part of the
  regularisation trick but this has been used previously for error
  (e.g. equation (1)).
\end{itemize}

@@

While the results in the paper provide some novel insights on the
topic, I believe the current version of the manuscript is far from the
standard required by Bayesian Analysis. Let me raise immediately my
main concern: on p.4 the author describes the generality of
measurements clusterisation as the first main question being explored,
which is answered via, quoting, "randomized numerical simulations that
exhibit clusterization more than 95\% of the time (see the code in
supplementary material)". However, in the version of the supplementary
materials I had access to there is no further mention of these
numerical simulations and results; only a link to a Github repository
is provided, which is however not functioning. Evidently, this issue
needs to be fixed before any future consideration can be given to the
manuscript. For example, a detailed description of the repeated
experiments could be provided either in the main article or in the
supplement, with tables indicating the various values of the
parameters for the simulations and the precise percentages
obtained. Also, the code needs to be publicly available for
reproducibility.

\subsection{Further comments}
I have a number of further comments in regards to the results and the overall presentation of
the manuscript.
\begin{itemize}
\item D-optimal design vs. statistical recovery rates: being more
  familiar with the statistical inverse problem literature than with
  D-optimal design, the clusterisation phenomenon demonstrated in this
  paper raises the question as to wether consistent recovery of the
  unknown (m in the notation of this paper) is possible under this
  choice. All the results I am aware of either assume white
  noise/equally spaced design (e.g. [1]) or measurement loctions
  sampled uni- formly at random (e.g. [2]). There are also some
  results with more general design under conditions on the
  fill-distance of the grid (e.g. [3]). All of these specifically
  prevent clusteri- sation, which is key to the statistical
  analysis. Hence, it is not clear to me even if D-optimal design
  should be pursued at all, if it indeed it leads to clusterisation as
  implied by the present 1paper. I think a discussion on this would be
  illuminating for the reader and help draw a connection to the
  broader statistical inverse problems literature.

@@

  Answer: It is impossible to visually discern the measurements, even
  when resolution is increased.

@@

  Answer: That is a great point! I added the followng discussion:

@@

@@

@@

@@

Answer: Thanks for this comment but I would rather keep this section
here since...

@@

\item Section 4: the conclusion reached at the end of the section on
  p.15 seem to imply that in the noiseless limit, repeating any
  measurement does not improve the objective criterion, if corre-
  lated error model are present. Is it clear however that adding any
  (non-reapted) measurement always strictly increases it? I.e., there
  could be situations where adding one measurement does not change the
  value of the optimised criterion? Also, can the conclusion drawn
  here be extended to the case of noisy observations σ > 0?

\item At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
  defined for $\sigma^2= 0$, but in the latter case, could not a Gaussian
  posterior measure be still defined as in Gaussian process
  regression, e.g.[4]

\item Theorem 1 is expressed in terms of the prior covariance matrix,
  which is a user specified quantity. Hence, it seems to me that all
  sort of design behaviour under the D-optimal crite- rior can occur
  by engineering the prior. It would perhaps be informative to study
  in more details some representative example of inverse problem and
  Gaussian prior. For example, the recovery of the initial condition
  with the ‘Matérn-like’ Gaussian prior laid out in the sup- plement
  are good candidates. For these it is still not clear how the
  conclusion from Theorem 1 translates into the Figure 1, also because
  I am unsure about the applicability of the result to point
  evaluations. In particular, how the conclusion that with m = 4
  measurements the D-optimal procedure will aim to ignore the third
  and fourth eigenvalue/function was drawn?
\end{itemize}

>>>>

\answer: I appreciate the many thoughtful comments by yourself, the
associate editor and the reviewers. I also appreciate the option to
improve the paper. Please see my answers below in blue.

\begin{description}
  \item[a] the Introduction is to be made accessible to the whole BA
    readership;
    \answer: I have considerably extended the introduction.

  \item[b] the motivation and the practical statistical implications are to be
    developed and discussed;
    \answer: I added a section on the implications in the introduction.

  \item[c] the contribution has to be better contextualized with
    respect to the literature in both inverse problems and D-optimal
    designs;
  \answer:

  \item[d] numerical evidence and code, mentioned in the paper but
    actually not provided, need to be made available;

    \answer: I added a section describing numerical experiments and
    their results. I also added documentation to the accompanying
    \href{https://github.com/yairdaon/OED}{repository} and made sure
    all scripts there run.

  \item[e] the mathematical derivations have to be watertight and some
    clarifications are needed in this respect (see Report of Referee
    3).

    \answer: I am not sure which specific comment of Reviewer 3 this
    refers to, please see my answers to Reviewer 3 below.
\end{description}

@@

\answer: I have added considerably to the introduction. See revised
manuscript and diff file.

@@

However, I feel changing the assumption from independent to correlated
measurements is quite artificial: people tweak a data generation
process assumption to solve an optimal design problem, and the author
just takes the assumption as it is.

\answer: If I understand correctly --- I completely agree. I added
text emphasizing that I do not endorse using correlated errors.

Since the author proves the clusterization issue under the independent
measurement assumption, I expect the author to go deeper and explain
why an ordinary data generation assumption + a common optimal design
leads to a “suboptimal” design?
\answer: Aagain, my apologies for the misunderstanding.

Not just by the mathematical proof, but from an information theory
point of view. For example, does the redundancy points mean additional
points does not reduce the variance of the estimation at all? If
that’s the case, how does correlated measurement assumption “help”
with it?

\answer: Bla

@@

\answer: I appreciate the comment, I fixed the relevant line in the
manuscript.

@@

If one looks at, for example, Fedorov (1996) and Nyberg et
al (2012), it is an undesirable property because they are interested
in sensor location, and two or more sensors cannot be in the same
location. However, there are many experiments where repeated
measurements are possible, and are therefore optimal

\answer: I believe that even in an application where repeated
experiments are allowed they may be unintuitive. For example, in a
medical imaging application --- taking a repeated slice in an MRI scan
might seem suboptimal to some. Of course, there is no argument
regarding the optimality give a model, which, as you mention might be
wrong.

(if the model is correct, which it won’t be).

\answer: I believe clusterization may imply a model is somehow in
adequate, see text below.

An example, is a chemical engineering experiment
where one wishes to learn the relationship between a series of
variables and yield of a chemical reaction. The experiment will
involve specifying the variables, and measuring the yield (the
observation) after a specified period of time. This process is then
duplicated with a potentially different set of variables.  It is
perfectly possible to have repeated variables.

\answer: I couldn't agree more. Indeed, the suboptimality of clustered
design is a matter of personal opinion; once a design criterion and a
model are agreed upon there can be no disagreement.

@@

\answer: Thanks for this comment. I considerably expanded the
introduction and delegated many proofs to the Supplementary. I also
added a section on practical implications, reproduced below.

@@

\answer: To a certain extend I agree: indeed, the impossibility of
taking identical measurements is indeed natural. However, this is not
encoded in the mathematical formulation of the problem. One can then
take one of two approaches: either accept what the mathematics
suggest, even if it counterintuitive, or find ways to circumvent the
mathematics because the result seems non-intuitive.

@@

\answer: I do not support imposing correlations in the paper, I just
used correlations to show how my framework can be used to rigorously
prove what is widely accepted. Other referees also understood that the
paper endorses imposing correlations and I accept that this is my
error in writing. I added a clause clearing the issue.

\subsection{Minor comments}
\begin{itemize}
\item The author uses two different forms of asterisk: one for
  optimality of the design (e.g. Definition 3) and one for the
  adjoint. Suggest changing the one for optimality.

  \answer: I changed $\obs^\star$ to $\opt$.

\item I do not like the use of the term measurements as used in the
  paper. The measurements are the $\data$’s since these are a result
  of measuring a physical quantity. However, the paper refers to the
  quantities that are controlled as the measurements.

  \answer: I was not able to find a better term to replace
  "measurements", so I stuck to it.

\item Proof of Proposition 12 uses $\eps$ as part of the
  regularisation trick but this has been used previously for error
  (e.g. equation (1)).

  \answer: Absolutely correct, I changed it to $\nu$. You will not
  find the revised proof in the manuscript since I moved it to the
  Supplementray.
\end{itemize}

@@

While the results in the paper provide some novel insights on the
topic, I believe the current version of the manuscript is far from the
standard required by Bayesian Analysis. Let me raise immediately my
main concern: on p.4 the author describes the generality of
measurements clusterisation as the first main question being explored,
which is answered via, quoting, "randomized numerical simulations that
exhibit clusterization more than 95\% of the time (see the code in
supplementary material)". However, in the version of the supplementary
materials I had access to there is no further mention of these
numerical simulations and results; only a link to a Github repository
is provided, which is however not functioning.

\answer: I apologize for not properly maintainig the repository. It is
now functioning and well-documented.

Evidently, this issue needs to be fixed before any future
consideration can be given to the manuscript. For example, a detailed
description of the repeated experiments could be provided either in
the main article or in the supplement, with tables indicating the
various values of the parameters for the simulations and the precise
percentages obtained. Also, the code needs to be publicly available
for reproducibility.

\answer: I appreciate the suggestion. I added a section on numerical
simulations to the main atricle. It is reproduced below.

\subsection{Further comments}
I have a number of further comments in regards to the results and the
overall presentation of the manuscript.

\begin{itemize}
\item D-optimal design vs. statistical recovery rates: being more
  familiar with the statistical inverse problem literature than with
  D-optimal design, the clusterisation phenomenon demonstrated in this
  paper raises the question as to wether consistent recovery of the
  unknown (m in the notation of this paper) is possible under this
  choice. All the results I am aware of either assume white
  noise/equally spaced design (e.g. [1]) or measurement loctions
  sampled uniformly at random (e.g. [2]). There are also some
  results with more general design under conditions on the
  fill-distance of the grid (e.g. [3]). All of these specifically
  prevent clusterisation, which is key to the statistical
  analysis. Hence, it is not clear to me even if D-optimal design
  should be pursued at all, if it indeed it leads to clusterisation as
  implied by the present paper. I think a discussion on this would be
  illuminating for the reader and help draw a connection to the
  broader statistical inverse problems literature.

@@

  \answer: It is impossible to visually discern the measurements, even
  when resolution is increased. These measurements are not identical
  numerically, but they are identical to five decimal digits. Since I
  find these points via optimization over point location in
  $\Omega=[0,1]$ I view an error $<10^{-5}$ as reasonable.

@@

  \answer: That is a great point!

@@

  \answer: I added relevant simulations to a new section Numerical
  Experiments.

@@

  \answer: I removed the statement of Theorem 1 from the introudction.

@@

  \answer: Thank you for the comment, I completely agree. Optimal
  designs are now denoted $\opt$.

@@

  \answer: I am glad you view this part as unnecessary, but I feel the
  model might feel too abstract to some practitioners. Thus, I choose
  to keep this part to avoid complaints from other readers.

@@

  \answer: Basically I meant that its eigenvalues decrease
  "quickly". I understand this is not a commot or well-defined term so
  I removed the relevant sentence.

\item Section 4: the conclusion reached at the end of the section on
  p.15 seem to imply that in the noiseless limit, repeating any
  measurement does not improve the objective criterion, if correlated
  error model are present. Is it clear however that adding any
  (non-reapted) measurement always strictly increases it? I.e., there
  could be situations where adding one measurement does not change the
  value of the optimised criterion?

  \answer: Great comment! Such cases are possible but I view them as
  pathological. See short paragraph below.

  Also, can the conclusion drawn here be extended to the
  case of noisy observations $\sigma > 0$?

  \answer: I do not expect this conclusion to hold: if noise is
  present, any measurement will result in some information gain and an
  increase in design criterion.

\item At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
  defined for $\sigma^2= 0$, but in the latter case, could not a Gaussian
  posterior measure be still defined as in Gaussian process
  regression, e.g.[4].

  \answer: I believe a posterior can be defined, and so a KL
  divergence from said posterior to prior could be calculated and
  $\tar$ evaluated. However, taking $\sigma = 0$ in the main theorem
  implies $\tar \equiv \infty$. I believe this is because do not know
  how to mitigate this

\item Theorem 1 is expressed in terms of the prior covariance matrix,
  which is a user specified quantity. Hence, it seems to me that all
  sort of design behaviour under the D-optimal criterion can occur by
  engineering the prior. It would perhaps be informative to study in
  more details some representative example of inverse problem and
  Gaussian prior. For example, the recovery of the initial condition
  with the ‘Matérn-like’ Gaussian prior laid out in the supplement are
  good candidates.

  \answer: the Matern-like prior actually arises by using a
  Laplacian-like operator as a prior, see Rue and Lindgren \cite{}. So
  a study of this example (with and without correlated errors) is
  already included in the paper. I am afraid implementing another
  inverse problem is out of scope for the present paper since I was
  not able to install common packages for inverse problems on my machine.

  For these it is still not clear how the conclusion from Theorem 1
  translates into the Figure 1, also because I am unsure about the
  applicability of the result to point evaluations.

  \answer: Indeed, point evaluations $\delta_x$ are not in any
  function space I consider in the paper since my analysis is limited
  to Hilbert spaces. Point evaluations can be approximated well,
  though so in my opinion this is not a deal breaker. The only caveat
  is that upon approximating point evaluations e.g.~by $R(x;t) =
  t\mathbb{1}_{-t/2, t/2}(x)$ we will need to change the norm
  constraints to the norsomething different (i.e.~$\|R(\cdot;t\|$) but
  that does not change the analysis. I skim over this issue in the
  paper since I think it is confusing and adds little context.

  In particular, how the conclusion that with $m = 4$ measurements the
  D-optimal procedure will aim to ignore the third and fourth
  eigenvalue/function was drawn?

  \answer: I explain this in more detail in the paper. See text below.

\end{itemize}
